Rényi Entropy-Based Heterogeneity Detection in SAR Data

### 2025-02-11 Text/Articles/RENYI_ENTROPY_SAR_2025/R0-Renyi-entropy-sar.qmd
update paper with corrections

author:
  - id: 
    name: Janeth Alpala, Abraão D.\ C.\ Nascimento 
    affiliations:
      - name: Universidade Federal de Pernambuco
        department: Departamento de Estatística
        city: Recife 
        country: PE, Brazil
        postal-code: 18800
      - name: Unknown affiliation
    #orcid: 
    email: janeth.alpala@ufpe.br, abraao@de.ufpe.br
    #url: 
    #membership: "Member, IEEE"
    attributes:
      corresponding: true
    #photo: 
    #bio: |
    #  Use `IEEEbiography`  with figure as  option and
    #  the author name as the argument followed by the biography text.
  - name: Alejandro C.\ Frery
    affiliations:
      - name: Victoria University of Wellington
        department: School of Mathematics and Statistics
        city: Wellington 
        country: New Zealand
        postal-code: 18800
      - name: Unknown affiliation
    #orcid: 
    email: alejandro.frery@vuw.ac.nz
    # bio: |
    #   Use `IEEEbiographynophoto` and the author name
    #   as the argument followed by the biography text.
    # note: "Template created June 23, 2023; revised `r format(Sys.Date(),format='%B %d, %Y')`."
abstract: |
 We propose a test statistic for identifying roughness characteristics in synthetic aperture radar (SAR) intensity data using Rényi entropy. 
 Homogeneous regions, where speckle is fully developed, are modeled by the Gamma distribution, while heterogeneous areas, associated with surface roughness, follow the $\mathcal{G}^0$ distribution. 
 The test relies on a non-parametric estimator of Rényi entropy to distinguish between these two types of regions. 
 By applying it to real SAR images, we generate $p$-value maps to assess the homogeneity hypothesis. 
 Experimental results show that the Rényi-based approach improves heterogeneity detection compared to previous methods using Shannon entropy.

keywords: [Gamma distribution, heterogeneity, SAR, Rényi entropy, hypothesis tests]
#funding: 
funding: 
  statement: "The `quarto-ieee` template is freely available under the MIT license on github: <https://github.com/dfolio/quarto-ieee>."
pageheader:
  left: Journal XXX, Month Year
  right: #'D. Folio:  A Sample Article Using quarto-ieee'

@@

Synthetic Aperture Radar (SAR) technology has become essential for a wide range of applications, including environmental monitoring and disaster management\ [@Yu2023;  @Mondini2021]. 
By operating independently of sunlight and under various weather conditions, SAR systems provide continuous, high-resolution imagery\ [@Zeng2020]. 
Despite these advantages, SAR data exhibit speckle, a noise-like interference effect that arises due to the coherent nature of the imaging process\ [@Argenti2013;@Choi2019].
In intensity format, speckle is typically non-Gaussian, and its presence complicates subsequent image analysis tasks.

The $\mathcal{G}^0$ distribution has shown to be an effective model for SAR intensity data,  as it captures different levels of roughness.
A special case is the Gamma distribution,which arises when speckle is fully developed, indicating homogeneous regions.
Although these models provide flexible frameworks for describing SAR intensities, selecting the appropriate model can be challenging for two main reasons: the small samples used in practical applications  and the inherent difficulties associated with parameter estimation. 
These challenges complicate model selection and highlight the need to explore alternative statistical approaches.

@@

The rest of this article is organized as follows.
Section \ref{sec:pre} presents an overview of statistical models for SAR intensity data and introduces Rényi entropy.
In Section \ref{sec:met}, we describe the proposed hypothesis test and the formulation of the test statistic. 
Section \ref{sec:app} evaluates the test's performance using real SAR data.
Finally, Section \ref{sec:conclusion} presents the conclusions.

@@

The main models considered for SAR intensity data are the $\Gamma_{\text{SAR}}$ distribution, which is suitable for fully developed speckle, and the $G_I^0$ distribution, which is able to describe roughness\ [@Frery1997]. These distributions are characterized by the following probability density functions (pdfs):

\begin{align}
	f_Z(z;L, \mu\mid \Gamma_{\text{SAR}})&=\frac{L^L}{\Gamma(L)\mu^L}z^{L-1}\exp\left\{-Lz/\mu\right\} \mathbbm 1_{\mathbbm R_+}(z),\label{E:gamma1}
\end{align}
\begin{multline}
\label{E:gi01}
	f_Z\big(z; \mu, \alpha, L\mid G_I^0\big) = \frac{L^L\Gamma(L-\alpha)}{\big[-\mu(\alpha+1)\big]^{\alpha}\Gamma(-\alpha)\Gamma(L)}\\ \frac{z^{L-1}}{\big[-\mu(\alpha+1)+Lz\big]^{L-\alpha}}\mathbbm 1_{\mathbbm R_+}(z),
\end{multline}

where $\mu > 0$ is the mean,
$\alpha < 0$ measures the roughness, $L \geq 1$ is the number of
looks, $\Gamma(\cdot)$ is the gamma function, and
$\mathbbm 1_{A}(z)$ is the indicator function of the set $A$.

As demonstrated by\ [@Frery1997], the $\Gamma_{\text{SAR}}$  model is a particular case of the $G_I^0$ distribution. Specifically, for a given $\mu$ fixed,
$$
f_Z\big(z; \mu, \alpha, L\mid G_I^0\big)
\longrightarrow 
f_Z(z;L, \mu\mid \Gamma_{\text{SAR}}) 
$$
when $\alpha\to-\infty$.

@@

Introduced by Alfréd Rényi in 1961\ [@renyi1961measures], the Rényi entropy generalizes several entropy measures, including the Shannon entropy. For a continuous random variable $Z$ with pdf $f(z)$, the Rényi entropy of order $\lambda$ (where $\lambda > 0$ and $\lambda \neq 1$) is defined as:

\begin{align}
\label{E:entropy2}
H_\lambda(Z) = \frac{1}{1 - \lambda} \ln \int_{-\infty}^{\infty} [f(z)]^\lambda \, dz.
\end{align}

Using this definition, we obtain the Rényi entropy expressions for the $\Gamma_{\text{SAR}}$ and $G_I^0$ distributions, respectively:

\begin{multline}
\label{eq-HGammaSAR}
H_\lambda\bigl(\Gamma_{\mathrm{SAR}}(L, \mu)\bigr)
= 
\ln \mu - \ln L + \frac{1}{1-\lambda}
\Bigl[
  -\lambda\,\ln\Gamma(L) \\  + \ln\Gamma\bigl(\lambda(L-1)+1\bigr)  - \bigl(\lambda(L-1)+1\bigr)\,\ln\lambda
\Bigr]
\end{multline}
and
\begin{multline}\label{eq-HGI0}
H_\lambda\bigl(G_I^0(\mu, \alpha, L)\bigr)=\ln\mu-\ln L +\ln(-1 - \alpha)\\
+ \frac{1}{\,1 - \lambda\,}
\Bigl[
   \lambda\bigl(
      \ln\Gamma(L - \alpha)
      -\ln\Gamma(-\alpha)
      -\ln\Gamma(L)
   \bigr)\\
   +\ln\Gamma\bigl(\lambda(L - 1) + 1\bigr)
   +\ln\Gamma\bigl(\lambda(-\alpha + 1) - 1\bigr)\\
   -\ln\Gamma\bigl(\lambda(L - \alpha)\bigr)
\Bigr].
\end{multline}

@fig-plot presents the Rényi entropy of $G_I^0$ as a function of $\mu$ for different $\alpha$ values. As $\alpha \to -\infty$, it approaches the Rényi entropy of $\Gamma_{\text{SAR}}$, which is consistent with the fact that $\Gamma_{\text{SAR}}$ is a limiting case of the $G_I^0$ model.

```{r fig-plot, echo=FALSE, message=FALSE, warning=FALSE, fig.fullwidth = TRUE, out.width="45%",  fig.pos="hbt", fig.cap="$H_{\\lambda}(G_I^0)$ converges to the $H_{\\lambda}(\\Gamma_{\\text{SAR}})$ when $\\alpha\\to-\\infty$, with $L=8$ and $\\lambda=0.8$.", fig.width=5, fig.height=4.0}

@@

Several authors have studied the nonparametric estimation of $H(Z)$, utilizing estimators that rely on spacings, which are the differences between order statistics\ [@vasicek1976test; @Ebrahimi1994; @Wieczorkowski1999;@AlOmari2019].
In 2024, Al-Labadi et al.\ [@AlLabadi2024] introduced a nonparametric estimator for Rényi entropy using this approach.

Let $\{Z_1, Z_2, \dots, Z_n\}$ be a sample from a distribution $F$, and denote its order statistics by $Z_{(1)} \leq Z_{(2)} \leq \dots \leq Z_{(n)}$. 
The density estimator based on $m$-spacing, is defined as:
$$
f_n(Z_{(i)}) = \frac{c_i m / n}{Z_{(i+m)} - Z_{(i-m)}},
$$

where $Z_{(i-m)} = Z_{(1)}$ when $i \leq m$ and $Z_{(i+m)} = Z_{(n)}$ if $i \geq n - m$. The coefficient $c_i$ is given by:
$$
c_i = 
\begin{cases}
\frac{m + i - 1}{m}, & \text{if } 1 \leq i \leq m, \\[6pt]
2, & \text{if } m+1 \leq i \leq n - m, \\[6pt]
\frac{n + m - i}{m}, & \text{if } n - m + 1 \leq i \leq n.
\end{cases}
$$
Following Vasicek [@vasicek1976test] and Ebrahimi et al. [@Ebrahimi1994] for Shannon entropy estimation, and using the $m$-spacing density estimator, Rényi entropy can be estimated as:
\begin{align}
\label{eq:est_R}
\widehat{H}_\lambda(\bm{Z}) = \frac{1}{1 - \lambda} \ln \left[\frac{1}{n} \sum_{i=1}^{n} \left( \frac{c_i m / n}{Z_{(i+m)} - Z_{(i-m)}} \right)^{\lambda - 1} \right].
\end{align}
This estimator is asymptotically consistent, i.e., it converges in
probability to the true value when $m,n\rightarrow\infty$ and
$m/n\rightarrow0$. 
We choose to use the heuristic formula for spacing, $m=\left[\sqrt{n}+0.5\right]$.

@@

We aim to determine the optimal order $\lambda$ for the Rényi entropy estimator, specifically for a sample size of $n=49$.
To identify this optimal value, we analyze both the Mean Squared Error (MSE) and Bias of the estimator across different values of $\lambda$. Lower MSE and Bias indicate better performance of the estimator in approximating the true entropy.

Based on the results, we find that the optimal value is $\lambda = 0.9$, as it minimizes the MSE while maintaining a low Bias. This choice is consistent with other values of $L > 1$. However, for $L = 1$, the optimal $\lambda$ tends to be higher (e.g., $\lambda = 3$) to achieve good results. @fig-plotf illustrates the case for $L = 5$.

```{r fig-plotf, echo=FALSE, message=FALSE, warning=FALSE, fig.fullwidth = TRUE, out.width="45%",  fig.pos="H", fig.cap="Bias and MSE as a function of $\\lambda$, with $n=49$, $L=5$.", fig.width=8, fig.height=6}

@@

Following previous studies by\ [@Frery2024] and\ [@Alpala2024], we also refine the non-parametric entropy estimator $\widehat{H}_{\lambda}$ using the bootstrap technique and apply an improved version:
$$
\widetilde{H}_{\lambda} = 2\widehat{\theta}(\bm{Z}) - \frac{1}{B} \sum_{b=1}^{B} \widehat{\theta}_b(\bm{Z}^{(b)}).
$$
A Monte Carlo study with $1000$ samples from $\Gamma_{\text{SAR}}$ ($\mu = 1$, $L=5$, $\lambda = 0.9$) confirms that this approach, using $B = 200$ replications, reduces bias and MSE for small sample sizes, as shown in @fig-Plot_bias_msef3.

@@

```{r fig-Plot_bias_msef3, echo=FALSE, message=FALSE, warning=FALSE, fig.fullwidth = TRUE, out.width="45%", fig.pos="H",  fig.cap="Bias and MSE of the Rényi entropy estimator for the $\\Gamma_{\\text{SAR}}$, with $\\lambda=0.9$, $\\mu=1$ and $L=5$.", fig.width=8, fig.height=7}

@@

Our approach relies on constructing a test statistic that compares a non-parametric estimate of the Rényi entropy to the theoretical value of the Rényi entropy under the null hypothesis (i.e., a fully developed speckle scenario modeled by $\Gamma_{\text{SAR}}$). 
We aim to test the following hypotheses: 
$$
\begin{cases}\mathcal{H}_0: \widehat{H}^*= H_{\lambda}(\Gamma_{\text{SAR}})\\ 
  \mathcal{H}_1:\widehat{H}^*= H_{\lambda}(G_I^0).\end{cases}
$$

@@

Assume we have an estimator $\widehat{H_{\lambda}}$ for the Rényi entropy of an arbitrary model. In the case of the $\Gamma_{\text{SAR}}$ model, this estimator is expected to be close to the theoretical expression given in Equation \eqref{eq-HGammaSAR}. 
Considering that $L\geq1$ is known, we define the following test statistic:
\begin{multline}
\label{eq-test}
S(\bm{Z}; L) = \widehat{H} - \bigl\{\ln \widehat{\mu} - \ln L + \frac{1}{1-\lambda}
\bigl[-\lambda\,\ln\Gamma(L) \\  
+ \ln\Gamma\bigl(\lambda(L-1)+1\bigr)  
- \bigl(\lambda(L-1)+1\bigr)\,\ln\lambda
\bigr]\bigr\}.
\end{multline}

When the null hypothesis holds, this test statistic is expected to be close to zero.

For illustrative purposes, @fig-density_entropyR provide the empirical density of $S(\bm{Z}; L)$  for the $\Gamma_{\text{SAR}}$ model, with $\lambda=0.9$ and $L \in \{5,18\}$.
For subsequent simulations we use the $\widetilde{H_{\lambda}}$ estimator.

@@

Therefore, the $p$-value is computed as $2\Phi(-\mid \varepsilon \mid)$,
where $\Phi$ is the standard Gaussian cumulative distribution function, and $\varepsilon$ is given by:  
$$
\varepsilon = \frac{\widetilde{H_i} - \widehat{H_i}(\Gamma_{\text{SAR}})}{\widehat{\sigma}_{i}}.
$$
An important aspect in hypothesis testing is ensuring that the test maintains the expected Type I error rate (size) while achieving high power (sensitivity to departures from $\mathcal{H}_0$).
To assess these properties, we performed $1000$ simulations at significance levels of $1\%$, $5\%$, and $10\%$, evaluating the test under the null hypothesis ($\Gamma_{\text{SAR}}$ distribution) with varying sample sizes, values of $L$, and $\mu=1$. 
The observed Type I error rates aligned with the nominal values, confirming the test’s validity.

Additionally, we examined the power under the alternative hypothesis ($G_I^0$ distribution) with $\mu=1$, $\lambda=0.9$ and $\alpha=-2$.
As expected, power improves as the sample size and number of looks increase the effectiveness demonstrating of the test. The results are shown in Table \ref{tab:table_size_power}.

```{r Simulated_error_type_I, echo=FALSE, message=FALSE}
set.seed(1234567890, kind = "Mersenne-Twister")
# 
calculate_p_value <- function(test_statistic, mu_W, sigma_W, alpha_nominal) {
  epsilon <- test_statistic / sigma_W
  p_value <- 2 * (1 - pnorm(abs(epsilon)))

  return(p_value < alpha_nominal)  
}
R <- 1000
mu <- 1
B <- 100
lambda <- 0.9
L_values <- c(  5, 8, 18)
sample_sizes <- c( 25, 49, 81, 121)
alpha_nominals <- c(0.01, 0.05, 0.1)
results <- data.frame()
for (L in L_values) {
  for (alpha_nominal in alpha_nominals) {
    TestStatistics <- NULL
    mean_entropy <- numeric(length(sample_sizes))
    sd_entropy <- numeric(length(sample_sizes))

@@

      for (r in 1:R) {
        z <- gamma_sar_sample(L, mu, s)
        TestStat[r] <-bootstrap_renyi_entropy_estimator_v1(z, B, lambda) -((lambda * lgamma(L) - lgamma(lambda * (L - 1) + 1)+(lambda * (L - 1) + 1) * log(lambda)) / (lambda - 1) +log(mean(z))-log(L))

        #TestStat[r] <-bootstrap_al_omari_1_estimator(z,B) - (log(mean(z)) + (L - log(L) + lgamma(L) + (1 - L) * digamma(L)))
        #TestStat[r] <- bootstrap_correa_estimator(z, B) - (L - log(L) + lgamma(L) + (1 - L) * digamma(L))
      }
      mean_entropy[sample_sizes == s] <- mean(TestStat)
      sd_entropy[sample_sizes == s] <- sd(TestStat)

      TestStatistics <- rbind(TestStatistics, data.frame("L" = rep(L, R), "Sample_Size" = rep(s, R), "Test_Statistics" = TestStat))
    }

@@

    p_values <- apply(TestStatistics, 1, function(row) {
      calculate_p_value(row["Test_Statistics"], mu_W[sample_sizes == row["Sample_Size"]], sigma_W[sample_sizes == row["Sample_Size"]], alpha_nominal)
    })

    result <- data.frame("L" = TestStatistics$L, "Sample_Size" = TestStatistics$Sample_Size, "Alpha_Nominal" = alpha_nominal, "P_Value" = p_values)
    results <- rbind(results, result)
  }
}
save(results, file = "./Data/type_I_results.Rdata")

```{r Simulated_power, echo=FALSE, message=FALSE}
set.seed(1234567890, kind = "Mersenne-Twister")
calculate_p_value <- function(test_statistic, mu_W, sigma_W) {
  epsilon <- test_statistic  / sigma_W
  p_value <- 2 * (1 - pnorm(abs(epsilon)))

>>>>

author:
  - id: 
    name: Janeth Alpala, Abraão D.\ C.\ Nascimento 
    affiliations:
      - name: Universidade Federal de Pernambuco
        department: Departamento de Estatística
        city: Recife 
        country: PE, Brazil
        postal-code: 18800
      - name: Unknown affiliation
    #orcid: 
    email: janeth.alpala@ufpe.br, abraao@de.ufpe.br
    #url: 
    #membership: "Member, IEEE"
    attributes:
      corresponding: true
    #photo: 
    #bio: |
    #  Use `IEEEbiography`  with figure as  option and
    #  the author name as the argument followed by the biography text.
  - name: Alejandro C.\ Frery
    affiliations:
      - name: Victoria University of Wellington
        department: School of Mathematics and Statistics
        city: Wellington 
        country: New Zealand
        postal-code: 18800
      - name: Unknown affiliation
    #orcid: 
    email: alejandro.frery@vuw.ac.nz
    # bio: |
    #   Use `IEEEbiographynophoto` and the author name
    #   as the argument followed by the biography text.
    # note: "Template created June 23, 2023; revised `r format(Sys.Date(),format='%B %d, %Y')`."
abstract: |
  Synthetic aperture radar (SAR) systems have already been successfully used to solve remote sensing problems. A disadvantage of SAR images is the presence of speckle, which is fully developed in homogeneous areas and gamma-distributed in these scenes. In heterogeneous areas, the intensity values are $\mathcal{G}^0_I$-distributed. 
  In this way, the identification of roughness SAR areas (as opposed to homogeneous ones) is an important task. 
  In this work, we propose a family of hypothesis tests driven by an order parameter to identify roughness features in SAR intensity data using Rényi entropy. 
  In particular, we use a nonparametric estimator for the Rényi entropy and investigate some of its properties. 
  As a practical evaluation method, we develop $p$-value maps on which one can observe both (i) the heterogeneous evidence change per texture and (ii) the prediction of homogeneous and heterogeneous categories. 
  The results are in favor of the Rényi-based heterogeneity detector compared to the one based on Shannon entropy.
keywords: [Rényi entropy, Gamma distribution, heterogeneity, SAR,  hypothesis tests]
#funding: 
funding: 
  statement: "The `quarto-ieee` template is freely available under the MIT license on github: <https://github.com/dfolio/quarto-ieee>."
pageheader:
  left: Journal XXX, Month Year
  right: #'D. Folio:  A Sample Article Using quarto-ieee'

@@

Synthetic Aperture Radar (SAR) technology has become essential for a wide range of applications\ [@Yu2023;@Akbarizadeh2012;@Mondini2021]. It provides high-resolution data that is independent of sunlight and operates in a variety of weather conditions to facilitate global Earth monitoring\ [@Zeng2020].
Despite its advantages, the effective use of SAR data depends on understanding its statistical properties because this data is affected by speckle, a noise-like interference effect\ [@Argenti2013; @Choi2019]. In intensity format, speckle is typically non-Gaussian, and its presence complicates subsequent image analysis tasks.

The $\mathcal{G}^0$ distribution has been shown to be an effective model for SAR intensity data, as it captures different levels of roughness. A special case is the Gamma distribution, which arises when speckle is fully developed, indicating homogeneous regions. 
Although these models offer flexibility in describing SAR intensities, selecting the appropriate model can be challenging for two main reasons: the small sample sizes used in practical applications and the inherent difficulties associated with parameter estimation. 
These challenges complicate model selection and highlight the need to explore alternative statistical approaches.

@@

The remainder of this article is organized as follows.
Section\ \ref{sec:pre} outlines an overview of statistical models for SAR intensity data and introduces Rényi entropy.
Section\ \ref{sec:met} describes the proposed hypothesis test and the formulation of the test statistic. 
Section\ \ref{sec:app} evaluates the performance of the test using real SAR data.
Finally, the concluding remarks are drawn in Section\ \ref{sec:conclusion}.

@@

The main models considered for SAR intensity data are the $\Gamma_{\text{SAR}}$ distribution, which is suitable for fully developed speckle, and the $\mathcal{G}^0_I$ distribution, which is able to describe roughness\ [@Frery1997]. These distributions are characterized by the following probability density functions (pdfs):
\begin{align}
	f_{\Gamma_{\text{SAR}}}\bigl(z;L, \mu \bigr) 
    &= \frac{L^L}{\Gamma(L)\,\mu^L} z^{L-1} 
    \exp \biggl(-\frac{Lz}{\mu}\biggr)
    \mathbbm 1_{\mathbbm R_+}(z) \label{E:gamma1}
    \intertext{and}
    f_{\mathcal{G}^0_I}\bigl(z; \mu, \alpha, L \bigr) 
    &= \frac{L^L\,\Gamma(L-\alpha)}
    {\bigl[-\mu(\alpha+1)\bigr]^{\alpha} \Gamma(-\alpha)\,\Gamma(L)}
    \notag \\
    &\quad \times
    \frac{z^{L-1}}
    {\bigl[-\mu(\alpha+1)+Lz\bigr]^{L-\alpha}}
    \mathbbm 1_{\mathbbm R_+}(z), \label{E:gi01}
\end{align}
where $\mu > 0$ is the mean,
$\alpha < 0$ measures the roughness, $L \geq 1$ is the number of
looks, $\Gamma(\cdot)$ is the gamma function, and
$\mathbbm 1_{A}(z)$ is the indicator function of the set $A$. 
As demonstrated by\ [@Frery1997], the $\Gamma_{\text{SAR}}$  model is a particular case of the $\mathcal{G}^0_I$ distribution. Specifically, for a given $\mu$ fixed,
$$
f_{\mathcal{G}^0_I}\big(z; \mu, \alpha, L\big)
\longrightarrow 
f_{\Gamma_{\text{SAR}}}(z;L, \mu) 
$$
when $\alpha\to-\infty$.

@@

Introduced by Alfréd Rényi in 1961\ [@renyi1961measures], this measure generalizes several well-known entropies, including the Shannon entropy\ [@Ribeiro2021]. For a continuous random variable $Z$ with pdf $f(z)$, the Rényi entropy of order $\lambda$, with $\lambda > 0$ and $\lambda \neq 1$, is defined as:
\begin{align}
\label{E:entropy2}
H_\lambda(Z) = \frac{1}{1 - \lambda} \ln \int_{-\infty}^{\infty} [f(z)]^\lambda \, dz.
\end{align}
Using \eqref{E:entropy2}, we derive closed-form expressions for the Rényi entropy of the $\Gamma_{\mathrm{SAR}}$ and $\mathcal{G}^0_I$ distributions:
\begin{multline}
\label{eq-HGammaSAR}
H_\lambda\bigl(\Gamma_{\text{SAR}}(L, \mu)\bigr)
= 
\ln \mu - \ln L + \frac{1}{1-\lambda}
\Bigl[
  -\lambda\,\ln\Gamma(L) \\  + \ln\Gamma\bigl(\lambda(L-1)+1\bigr)  - \bigl(\lambda(L-1)+1\bigr)\,\ln\lambda
\Bigr]
\end{multline}
and
\begin{multline}\label{eq-HGI0}
H_\lambda\bigl(\mathcal{G}^0_I(\mu, \alpha, L)\bigr)=\ln\mu-\ln L +\ln(-1 - \alpha)\\
+ \frac{1}{\,1 - \lambda\,}
\Bigl[
   \lambda\bigl(
      \ln\Gamma(L - \alpha)
      -\ln\Gamma(-\alpha)
      -\ln\Gamma(L)
   \bigr)\\
   +\ln\Gamma\bigl(\lambda(L - 1) + 1\bigr)
   +\ln\Gamma\bigl(\lambda(-\alpha + 1) - 1\bigr)\\
   -\ln\Gamma\bigl(\lambda(L - \alpha)\bigr)
\Bigr].
\end{multline}
@fig-plot presents the Rényi entropy of $\mathcal{G}^0_I$ as a function of $\mu$ for different $\alpha$ values. As $\alpha \to -\infty$, it approaches the Rényi entropy of $\Gamma_{\text{SAR}}$, which is aligned with the fact that $\Gamma_{\text{SAR}}$ is a limiting case of the $\mathcal{G}^0_I$ model.

```{r fig-plot, echo=FALSE, message=FALSE, warning=FALSE, fig.fullwidth = TRUE, out.width="45%",  fig.pos="hbt", fig.cap="$H_{\\lambda}(\\mathcal{G}^0_I)$ converges to the $H_{\\lambda}(\\Gamma_{\\text{SAR}})$ when $\\alpha\\to-\\infty$, with $L=8$ and $\\lambda=0.8$.", fig.width=5, fig.height=4.0}

@@

Nonparametric estimation of $H(Z)$ has been widely studied using spacing-based estimators, which rely on differences between order statistics [@vasicek1976test; @Ebrahimi1994;@Wieczorkowski1999; @AlOmari2019]. Recently, Al-Labadi et al.\ [@AlLabadi2024] proposed a nonparametric estimator for Rényi entropy following this approach.

<!-- Let $\bm{Z} = (Z_1, Z_2, \dots, Z_n)$ be a random sample from $F(z)$ with order statistics $Z_{(1)} \leq Z_{(2)} \leq \dots \leq Z_{(n)}$. The $m$-spacing density estimator is defined as: -->
Let $\bm{Z}=(Z_1, Z_2,\ldots,Z_n)$ be a random sample from the distribution $F(z)$ and denote its order statistics by $Z_{(1)} \leq Z_{(2)} \leq \dots \leq Z_{(n)}$.
The density estimator based on $m$-spacing, is defined as:
$$
f_n(Z_{(i)}) = \frac{c_i m / n}{Z_{(i+m)} - Z_{(i-m)}},
$$
where $Z_{(i-m)} = Z_{(1)}$ when $i \leq m$ and $Z_{(i+m)} = Z_{(n)}$ if $i \geq n - m$. The coefficient $c_i$ is given by:
$$
c_i = 
\begin{cases}
\frac{m + i - 1}{m}, & \text{if } 1 \leq i \leq m, \\[6pt]
2, & \text{if } m+1 \leq i \leq n - m, \\[6pt]
\frac{n + m - i}{m}, & \text{if } n - m + 1 \leq i \leq n.
\end{cases}
$$

Following Vasicek [@vasicek1976test] and Ebrahimi et al. [@Ebrahimi1994] for Shannon entropy estimation, and using the $m$-spacing density estimator, the Rényi entropy can be estimated as:
\begin{align}
\label{eq:est_R}
\widehat{H}_\lambda(\bm{Z}) = \frac{1}{1 - \lambda} \ln \left[\frac{1}{n} \sum_{i=1}^{n} \left( \frac{c_i m / n}{Z_{(i+m)} - Z_{(i-m)}} \right)^{\lambda - 1} \right].
\end{align}
This estimator is asymptotically consistent, i.e., it converges in
probability to the true value when $m,n\rightarrow\infty$ and
$m/n\rightarrow0$. 
We choose to use the heuristic formula for spacing, $m=\left[\sqrt{n}+0.5\right]$.

@@

We aim to determine the optimal order $\lambda$ for the Rényi entropy estimator, specifically for a sample size of $n=49$.
To identify this optimal value, we analyze both the mean squared error (MSE) and bias of the estimator across different values of $\lambda$. Lower MSE and bias indicate better performance of the estimator in approximating the true entropy.

Based on the results, we find that the optimal value is $\lambda = 0.9$, as it minimizes the MSE while maintaining a low bias. This choice is consistent with other values of $L > 1$. However, for $L = 1$, the optimal $\lambda$ tends to be higher (e.g., $\lambda = 3$) to achieve good results. @fig-plotf illustrates the case for $L = 5$.

```{r fig-plotf, echo=FALSE, message=FALSE, warning=FALSE, fig.fullwidth = TRUE, out.width="45%",  fig.pos="hbt", fig.cap="Bias and MSE as a function of $\\lambda$, with $n=49$, $L=5$.", fig.width=8, fig.height=6}

@@

Following previous studies by\ [@Frery2024] and\ [@Alpala2024], we also refine the non-parametric entropy estimator $\widehat{H}_{\lambda}$ using the bootstrap technique and apply the improved version:
$$
\widetilde{H}_{\lambda} = 2\widehat{\theta}(\bm{Z}) - \frac{1}{B} \sum_{b=1}^{B} \widehat{\theta}_b(\bm{Z}^{(b)}),
$$
where $B$ is the number of new samples generated through resampling with replacement.

A Monte Carlo study with $1000$ samples from $\Gamma_{\text{SAR}}$ ($\mu = 1$, $L=5$, $\lambda = 0.9$) confirms that this approach, using $B = 200$ replications, reduces bias and MSE for small sample sizes, as shown in @fig-Plot_bias_msef3.

@@

```{r fig-Plot_bias_msef3, echo=FALSE, message=FALSE, warning=FALSE, fig.fullwidth = TRUE, out.width="45%", fig.pos="hbt",  fig.cap="Bias and MSE of the Rényi entropy estimator for the $\\Gamma_{\\text{SAR}}$, with $\\lambda=0.9$, $\\mu=1$ and $L=5$.", fig.width=8, fig.height=7}

@@

Our approach relies on constructing a test statistic that compares a non-parametric estimate of the Rényi entropy to the theoretical value of the Rényi entropy under the null hypothesis (i.e., a fully developed speckle scenario modeled by $\Gamma_{\text{SAR}}$). 
We aim to test the following hypotheses: 
$$
\begin{cases}\mathcal{H}_0: \widehat{H}^*= H_{\lambda}(\Gamma_{\text{SAR}})\\ 
  \mathcal{H}_1:\widehat{H}^*= H_{\lambda}(\mathcal{G}^0_I).\end{cases}
$$

@@

Assume we have an estimator $\widehat{H}_{\lambda}$ for the Rényi entropy of an arbitrary model. In the case of the $\Gamma_{\text{SAR}}$ model, this estimator is expected to be close to the theoretical expression given in Equation \eqref{eq-HGammaSAR}. 
Since $L\geq1$ is known, we define the test statistic as follows:
\begin{multline}
\label{eq-test}
S(\bm{Z}; L) = \widehat{H} - \bigl\{\ln \widehat{\mu} - \ln L + \frac{1}{1-\lambda}
\bigl[-\lambda\,\ln\Gamma(L) \\  
+ \ln\Gamma\bigl(\lambda(L-1)+1\bigr)  
- \bigl(\lambda(L-1)+1\bigr)\,\ln\lambda
\bigr]\bigr\}.
\end{multline}
When the null hypothesis holds, this test statistic is expected to be close to zero.

For illustrative purposes, @fig-density_entropyR provide the empirical density of $S(\bm{Z}; L)$  for the $\Gamma_{\text{SAR}}$ model, with $\lambda=0.9$ and $L \in \{5,18\}$.
For subsequent simulations we use the $\widetilde{H}_{\lambda}$ estimator.

@@

Therefore, the $p$-value is computed as $2\Phi(-\mid \varepsilon \mid)$,
where $\Phi$ is the standard Gaussian cumulative distribution function, and $\varepsilon$ is given by:  
$$
\varepsilon = \frac{\widetilde{H}_i - \widehat{H}_i(\Gamma_{\text{SAR}})}{\widehat{\sigma}_{i}}.
$$
An important aspect of hypothesis testing is ensuring that the test maintains the expected type I error rate (size) while achieving high power (sensitivity to departures from $\mathcal{H}_0$).
To assess these properties, we performed a Monte Carlo simulation with $1000$ replications at significance levels  $1\%$, $5\%$, and $10\%$, evaluating the test under the null hypothesis ($\Gamma_{\text{SAR}}$ distribution) and varying sample sizes, values of $L$, and $\mu=1$. 
The observed type I error rates aligned with the nominal values, confirming the test validity.

Additionally, we examined the power under the alternative hypothesis ($\mathcal{G}^0_I$ distribution) with $\mu=1$, $\lambda=0.9$ and $\alpha=-2$.
As expected, power improves as the sample size and number of looks increase the effectiveness demonstrating of the test. The results are shown in Table \ref{tab:table_size_power}.

```{r Simulated_error_type_I, echo=FALSE, message=FALSE}
if (!file.exists("./Data/type_I_results.Rdata")) {

  message("AFile type_I_results.Rdata not found. Generating results...")

  set.seed(1234567890, kind = "Mersenne-Twister")

  calculate_p_value <- function(test_statistic, mu_W, sigma_W, alpha_nominal) {
    epsilon <- test_statistic / sigma_W
    p_value <- 2 * (1 - pnorm(abs(epsilon)))
    return(p_value < alpha_nominal)  
  }

  R <- 1000
  mu <- 1
  B <- 100
  lambda <- 0.9
  L_values <- c(5, 8, 18)
  sample_sizes <- c(25, 49, 81, 121)
  alpha_nominals <- c(0.01, 0.05, 0.1)
  results <- data.frame()

  for (L in L_values) {
    for (alpha_nominal in alpha_nominals) {
      TestStatistics <- NULL
      mean_entropy <- numeric(length(sample_sizes))
      sd_entropy <- numeric(length(sample_sizes))

@@

        for (r in 1:R) {
          z <- gamma_sar_sample(L, mu, s)
          TestStat[r] <- bootstrap_renyi_entropy_estimator_v1(z, B, lambda) - (
            (lambda * lgamma(L) - lgamma(lambda * (L - 1) + 1) + 
             (lambda * (L - 1) + 1) * log(lambda)) / (lambda - 1) +
            log(mean(z)) - log(L)
          )
        }

        TestStatistics <- rbind(
          TestStatistics,
          data.frame(
            "L" = rep(L, R),
            "Sample_Size" = rep(s, R),
            "Test_Statistics" = TestStat
          )
        )
      }

@@

      p_values <- apply(TestStatistics, 1, function(row) {
        calculate_p_value(
          row["Test_Statistics"],
          mu_W[sample_sizes == row["Sample_Size"]],
          sigma_W[sample_sizes == row["Sample_Size"]],
          alpha_nominal
        )
      })

      result <- data.frame(
        "L" = TestStatistics$L,
        "Sample_Size" = TestStatistics$Sample_Size,
        "Alpha_Nominal" = alpha_nominal,
        "P_Value" = p_values
      )
      results <- rbind(results, result)
    }
  }

  save(results, file = "./Data/type_I_results.Rdata")

} else {

  message("File type_I_results.Rdata found. Generating results...")
  load("./Data/type_I_results.Rdata")

}
```

```{r Simulated_power, echo=FALSE, message=FALSE}

if (!file.exists("./Data/results_power.Rdata")) {

  message("File type_I_results.Rdata not found. Generating results...")

  set.seed(1234567890, kind = "Mersenne-Twister")

  calculate_p_value <- function(test_statistic, mu_W, sigma_W) {
    epsilon <- test_statistic  / sigma_W
    p_value <- 2 * (1 - pnorm(abs(epsilon)))
    return(p_value)
  }
